\titleNameLong{}

### 2015-12-08 FinalProject/Tex/paper/paper.tex
fixed line breaks on title
split dynamic section in two, finished first half

The remainder of this survey is organized as follows. Section
\ref{sec:architecture} provides a brief architectural overview of GPUs. Section
\ref{sec:challenges} elaborates on the challenges specific to GPGPU
applications. Section \ref{sec:dynamic} discusses various dynamic approaches to
scheduling and warp configurations. Section \ref{sec:memory} talks specifically
about reducing memory contention. Section \ref{sec:software} is a discussion of
a few software and compiler solutions to the challenges of GPGPU. Section
\ref{sec:conclusions} concludes the survey.

@@

\section{Dynamic Warp Scheduling and Hardware Tuning} \label{sec:dynamic}
\subsection{Assist warps (CABA)}
This section left as an exercise to the reader
\subsection{Coordinated warp scheduling (CAWA)}
This section left as an exercise to the reader
\subsection{Dynamic Thread Block Launch (DTBL)}
This section left as an exercise to the reader
\subsection{Equalizer}
This section left as an exercise to the reader
\subsection{Fine-grain Hardware Worklists (HWWL)}
This section left as an exercise to the reader
\subsection{Variable Warp Size (VWS)}
This section left as an exercise to the reader

\section{Reducing Memory Contention} \label{sec:memory}
\subsection{Warped Compression}
This section left as an exercise to the reader
\subsection{Managing Concurrency in Heterogeneous (CPU/GPU NoC contention)}
This section left as an exercise to the reader

\section{Compiler/SW-only design} \label{sec:software}
\subsection{Divergence Management}
This section left as an exercise to the reader
\subsection{Locality Aware Mapping}
This section left as an exercise to the reader
\subsection{(x) WeakBehaviorProgrammingAssumptions}
This section left as an exercise to the reader

>>>>

The remainder of this survey is organized as follows. Section
\ref{sec:architecture} provides a brief architectural overview of GPUs. Section
\ref{sec:challenges} elaborates on the challenges specific to GPGPU
applications. Section \ref{sec:dynamic} discusses spawning granular helpers to
increase performance. Section \ref{sec:scheduling} deals with more efficient
scheduling within and among streaming multiprocessors. Section \ref{sec:memory}
talks specifically about reducing memory contention. Section \ref{sec:software}
is a discussion of a few software and compiler solutions to the challenges of
GPGPU. Section
\ref{sec:conclusions} concludes the survey.

@@

\section{Dynamically-Launched Helpers} \label{sec:dynamic}
\subsection{Assist warps}
As we have seen, GPUs are capable of supporting thousands of threads, but
bottlenecks in heterogeneous applications limit performance. \cite{AssistWarps}
present the idea of using ``assist warps'' in their Core-Assisted Bottleneck
Acceleration (CABA) scheme. These special warps are generated by dedicated
hardware and are used as helper threads to increase performance and efficiency.
Assist warps share the same context as the parent warp and uses some of the
available registers, since investigation shows that 24\% of the register file is
unallocated on average.

One primary use for assist warps is memory compression. Instead of using
dedicated hardware for compressing and decompressing data from memory, this
approach uses existing hardware that is not being utilized. Another advantage of
not using dedicated hardware is the ability to change compression algorithms
flexibly, since different applications have varying access patterns that are
suited for different compression algorithms. Another use for assist warps is in
memoization, in which frequently- computed values from main memory are cached
and used as lookup tables, reducing the number of loads and stores required. One
final use suggested by the authors is using an idle memory pipeline to prefetch
values during computation to reduce the memory pressure.

CABA contains a mix of hardware and software. While software-only or hardware-
only options are possible, the former induces high overhead and complicates
communication between primary and helper warps, and the latter makes having a
flexible interface more difficult. The combination suggested by the authors
dynamically inserts instructions into the execution stream in the form of assist
warps. This requires stores of commonly used assist warps, a controller to
trigger and manage assist warp execution, and an assist warp buffer inside the
instruction buffer that holds triggered warps until they are scheduled. Helper
threads share context with the parent warp, and register values in helper
threads are not preserved between invocations, reducing the overhead of this
technique and does not require any additional resources.

The authors report a reduction in memory bandwidth of 2.1x, resulting in an
increase in performance of 41.7\%. This is only 2.8\% less than the idea case in
which there would be no overhead of assist warps. These helper threads also
reduced the overall system energy by 22.2\%, which is 4.0\% over the ideal
effect. Power consumption is also increased by 2.9\%, but this is a small price
to pay for the performance increases.

\subsection{Dynamic Thread Block Launch}
The authors of Dynamic Thread Block Launch (DTBL) suggest a somewhat similar
approach, this time launching thread blocks dynamically instead of assist warps
\cite{DynamicThreadBlockLaunch}. The motivation behind this technique is that
while emerging general-purpose applications do not have simple, regular
structure that is easy to exploit at the highest level, they still do contain
``pockets'' of parallelism at smaller levels. The authors explain that these
pockets can be exploited with existing architecture using device-side nested
kernel launching, but at a high cost. For example, CUDA supports parent kernels
launching child kernels through device-side API calls. These child kernels can
start at any time after being launched by their parent. This method does exploit
pockets of parallelism, but at a high cost. On average, this CUDA method results
in a 36.1\% decrease in performance, and also consumes significant amounts of
memory by storing pending child kernels.

The problems with existing implementations boil down to four characteristics:
\begin{description}
\item [High kernel density] Kernel launches are expensive, and typical
applications currently require many kernels. For example, a typical breadth-
first search requires about 3,000 device kernel launches, resulting in a high
overhead and memory footprint.
\item [Low compute intensity] While many kernels are launched using these
methods, they are usually fine-grained and do not contain high amounts of
parallelism. The average number of threads in these kernels is around 40, which
is close to the warp size and could be better served by a more fine-grained
approach.
\item [Workload similarity] The operations performed by each dynamically
launched kernel are usually similar, but contain different levels of parallelism
due to differing configurations or data.
\item [Low concurrency/scheduling efficiency] Due to the low compute intensity
of the device kernels, multiple kernels should be able to execute in parallel.
However, GPU architectures place limits on the number of different kernels that
can execute concurrently, and since each kernel has a relatively small number of
warps within it, the SMs are underutilized and struggle to hide memory
latencies.
\end{description}

DTBL is a more fine-grained approach. Instead of launching entire kernels,
individual thread blocks are created and added to the list of existing thread
blocks within a kernel. These thread blocks are coalesced into an aggregated
kernel and executed together. The existing GPU microarchitecture is extended to
support these additional thread blocks by adding data structures to keep track
of dynamically formed groups of thread blocks and associating them with kernels.
Fortunately, this organization is transparent to much of the rest of the
architecture, including the warp schedulers and control divergence mechanism.

The extra data structures of this technique only contribute an area overhead of
0.5\%. Additionally, the authors report a 1.21x performance increase over
original GPU implementations and 1.40x over implementations using existing
device-kernel launches.

\section{Dynamic Scheduling} \label{sec:scheduling}
\subsection{Coordinated warp scheduling (CAWA)}

\subsection{Variable Warp Size (VWS)}
This section left as an exercise to the reader

\subsection{Fine-grain Hardware Worklists (HWWL)}
This section left as an exercise to the reader

\subsection{Equalizer}
This section left as an exercise to the reader

\section{Reducing Memory Contention} \label{sec:memory}
\subsection{Warped Compression}
This section left as an exercise to the reader

\subsection{Managing Concurrency in Heterogeneous (CPU/GPU NoC contention)}
This section left as an exercise to the reader

\section{Compiler/SW-only design} \label{sec:software}
\subsection{Divergence Management}
This section left as an exercise to the reader

\subsection{Locality Aware Mapping}
This section left as an exercise to the reader

\subsection{(x) WeakBehaviorProgrammingAssumptions}
This section left as an exercise to the reader
